Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-12-09 vesicle_code/docs/permeableVes/main.tex
yny

\dropcap{W}ater balance is an essential process in the adaptation of a living cell to its environment \cite{AlbertsMolecularBiology}.  For example, erythrocytes demonstrate efficient exchange of water to participate in efficient body water homeostasis \cite{SugieIntaglietta2018_AmJPhysiolHCP}. Another example is the regulation of cell volume and the cell membrane tension in cell migration \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS}. The cell membrane is permeable to water and water-soluble molecules \cite{Dick1964_JTB,FettiplaceHaydon1980_PhysRev,DeamerBramhall1986_ChemPhysLipids,Grafmueller2019_ABLS}. In the presence of osmolarity (a contrast in the concentration of a solute impermeable to the membrane), water flows in or out of the membrane to adjust the interior cellular volume. In the red blood cell membrane, aquaporin-1 (AQP1) channels give rise to a much higher water permeability of $\sim 1.8\times 10^{-2}$ cm/s in erythrocytes \cite{YangMaVerkman2001_JBC} while the intrinsic water permeability of a lipid bilayer membrane is in the range of $10^{-4}-10^{-3}$ cm/s \cite{ThompsonHuang1966_ANYAS,FettiplaceHaydon1980_PhysRev,Grafmueller2019_ABLS,Dimova2020_GVB}.

Vesicles (self-enclosed lipid bilayer membranes) have been used as a model system to study reshaping, remodeling and scission of cell membranes due to osmotic stress \cite{OgleckaEtAl2014_eLife,CamposSaric2020_bioRxiv,Dimova2020_GVB}. Under an osmotic shock, a vesicle may undergo oscillatory variation in its enclosed volume \cite{ChabanonHoLiedberg2017_BJ}. In these situations, the osmolarity is the main drive for water transport across the membrane. Thus most modeling works ignore the mechanical effects (capillary pressure, bending force and membrane tension) on vesicles under an osmotic stress. However, the osmotic water flux is inevitably coupled with the mechanical pressure jump across the membrane\cite{yao-mor2017}, which could become comparable to osmotic stress when the membrane deformation is large \cite{LiMoriSun2015_PRL,yao-mor2017}.

The intrinsic water permeability of a lipid bilayer membrane depends on the temperature, ph value in the solvent, and the hydration  energy of the membrane.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure}[htp]
  \ifTikz
  \input{figures/parabolicComposite.tikz}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady-state shape of
  an impermeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. The results agree with Kaoui et
  al.~\cite{kao-bir-mis2009}. (b) The steady state shape of a
  semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates.}
\end{figure}

>>>>

\dropcap{W}ater balance is an essential process in the adaptation of a living cell to its environment \cite{AlbertsMolecularBiology}.  For example, erythrocytes demonstrate efficient exchange of water to participate in efficient body water homeostasis \cite{SugieIntaglietta2018_AmJPhysiolHCP}.  Another example is the simultaneous regulation of cell volume and the cell membrane tension during cell migration through trafficking of aquaporin proteins \cite{SaadounPapadopoulosWatanabeEtAl2005_JCS}.

The cell membrane is permeable to water and water-soluble molecules \cite{Dick1964_JTB,FettiplaceHaydon1980_PhysRev,DeamerBramhall1986_ChemPhysLipids,Grafmueller2019_ABLS}. In the presence of an osmotic gradient across the membrane (such as a contrast in the concentration of an impermeable solute), water flows in or out of the membrane to adjust the interior cellular volume. In the red blood cell membrane, aquaporin-1 (AQP1) channels give rise to a much higher water permeability of $\sim 1.8\times 10^{-2}$ cm/s in erythrocytes \cite{YangMaVerkman2001_JBC} while the intrinsic water permeability of a lipid bilayer membrane is in the range of $10^{-4}-10^{-3}$ cm/s \cite{ThompsonHuang1966_ANYAS,FettiplaceHaydon1980_PhysRev,Grafmueller2019_ABLS,Dimova2020_GVB}.

Vesicles (self-enclosed lipid bilayer membranes) have been used as a model system to study reshaping, remodeling and scission of cell membranes due to osmotic stress \cite{OgleckaEtAl2014_eLife,CamposSaric2020_bioRxiv,Dimova2020_GVB}. Under an osmotic shock, a vesicle may undergo oscillatory variation in its enclosed volume \cite{ChabanonHoLiedberg2017_BJ}. In these situations, the osmotic stress is the dominant driving force for water transport across the membrane. Thus most modeling works ignore the mechanical effects (capillary pressure, bending force and membrane tension) on vesicles under an osmotic shock. However, the osmotic water flux is inevitably coupled with the mechanical pressure jump across the membrane\cite{yao-mor2017}, which could become comparable to osmotic stress when the membrane deformation is large \cite{LiMoriSun2015_PRL,yao-mor2017}.

Without aquaporins or other membrane channels that promotes water permeability, the intrinsic water permeability is known to depend on the temperature \cite{}, solvent ph \cite{}, and the hydration  energy of the membrane \cite{}. In this work we focus on the intrinsic water permeability and study the hydrodynamics of a semipermeable membrane. Specifically we focus on the water permeability due to jump in normal membrane traction.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure}[htp]
%  \ifTikz
  \input{figures/parabolicComposite.tikz}
%  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady-state shape of
  an impermeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. The results agree with Kaoui et
  al.~\cite{kao-bir-mis2009}. (b) The steady state shape of a
  semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates.}
\end{figure}
